\subsection{(Candidate) Research Questions}

\begin{itemize}
\item {\bf RQ1.} Can defiant behaviour be identified from the modeling of the scenarios of the component and its environment (including the  other components in the global system)? 

\item {\bf RQ2.} Is it possible to resolve definant behaviours by creating a wrapper to satisfy the global requirements using existing defiant components?

\item {\bf RQ3.} If possible, how many changes are required to adapt the behaviour of a defiant component? Would such adapation benefit from having aspect-oriented wrappers?

\item {\bf RQ4.} Can the simulation of the self-adaptive aspect-oriented wrappers be implemented using aspect-oriented programming?

\item {\bf RQ5.} Can multiple aspect-oriented wrappers be applied to a more complex scenario correctly, and efficiently? 
\end{itemize}

\subsection{Evaluation}

\begin{itemize}
\item {\bf RQ1.} {\em Can defiant behaviour be identified from the scenarios of the component and its environment (including the  other components in the global system)? }

At the design time, we need to define the verifiable FSP properties that can be checked by the LTSA model checker.

At the runtime, we need to define the  monitoring conditions of the involved context variables based on the sensor readings, so that when the monitoring conditions are
triggering the defiant behaviours (traced to the execution of scenarios) are identified. We need to define necessary and sufficient conditions. 

\item {\bf RQ2.} {\em Is it possible to resolve definant behaviours by creating a wrapper to satisfy the global requirements using existing defiant components?}

At the design time, we need to model check the property that the composed scenarios do not violating the global requirements.

At the runtime, we need to ensure that the analysis can be done efficiently. Is it possible to use BDD to reduce the runtime requirements satisfaction down to a decision tree model that is efficient to compute? 

\item {\bf RQ3.}{\em  If possible, how many changes are required to adapt the behaviour of a defiant component? Would such adapation benefit from having aspect-oriented wrappers?}

At the design time, we need to measure the number of changes introduced to the behaviour model of a defiant component, with or without aspect-oriented wrappers.

At the runtime, we need to justify the extra indirection due to the aspect invocation. 

\item {\bf RQ4.} {\em Can the simulation of the self-adaptive aspect-oriented wrappers be implemented using aspect-oriented programming?}

At the design time, we need to measure the number of changes introduced to the behaviour model of a defiant component, with or without aspect-oriented wrappers.

At the runtime, we need to justify the extra indirection due to the aspect invocation does not causing too much overhead.

\item {\bf RQ5.} {\em Can multiple aspect-oriented wrappers be applied to a more complex scenario correctly, and efficiently? }

At the design time, we need to check the adverse condition of ``feature interaction'' does not matter to the composition of aspect wrappers.

At the runtime, we need to justify enforce the reliability / resilience of execution of the aspect compositions. 
\end{itemize}

